### 2018-10-04 CO759-ExtendendFormulations/assignments/A1.tex
Problem 3

\begin{document}
%Problem 1
\section{}
\paragraph{}
Let a polytope $Q\subseteq \R^n$ and a linear map $p:\R^n \rightarrow \R^d$ be an extension of a polytope $P\subseteq \R^d$. Since $p$ is a linear map, there exists a matrix $A \in \R^{d\times n}$ such that $p(x) = Ax$ for all $x \in Q$.
\paragraph{}
Let $k \in \Z$ such that $0\leq k \leq d$. We will show the existence of an injective map $f$ from the faces of $P$ of dimension $k$ to the faces of $Q$ of dimension at least $k$. This suffices to show the desired result.
\paragraph{}
Let $F$ be a face of $P$ of dimension $k$ such that $F = \{x \in P: c^Tx = \beta\}$. We define the map $f$ so that
$$f(F) = \{x \in Q: (c^TA)x = \beta\}.$$
We need to show that $f$ is well-defined, and that it is injective. To this end we begin by showing that $f(F)$ is a face of $Q$. Let $x \in Q$. Then
$$(c^TA)x = c^T(Ax) \leq \beta$$
with the last inequality following since $p(x) = Ax$ is in $P$. Note that equality holds for all $x$ such that $Ax$ is in $F$. Hence $f(F)$ is a face of $Q$. Since $(Q,p)$ is an extension of $P$, for all $y \in F$ there exists $x \in Q$ such that $Ax = y$. Thus $F \subseteq p(f(F))$, and using that $p$ is a linear map we see that, 
$$\text{dim}(f(F)) \geq \dim F = k.$$
\paragraph{}
Now suppose that $F$ and $F'$ are $k$-dimensional faces of $P$ such that $f(F) = f(F')$. Suppose $F = \{x \in P: c^Tx = \beta\}$ and $F' = \{x \in P: (c')^Tx = \beta\}$. Let $y \in F$. Then there exists $x \in f(F)$ such that $Ax = y$. Since $f(F) = f(F')$,
$$\beta = (c')^TAx = (c')^Ty $$
and hence $y \in F$. Thus $F \subseteq F'$ and by a symmetric argument we can see that $F' \subseteq F$. Thus $F = F'$ and hence $f$ is an injective map, as desired. $\blacksquare$
%Problem 2
\section{}
\paragraph{}
The claim is \emph{false}. Let $P = \conv{\{(0,0),(1,0),(1,1),(0,1)\}}$. Then $P$ is a square, so $|\text{facets}(P)| = 4$. Let $Q = \conv{\{(0,0,0),(1,0,0), (1,1,0), (0,1,1)\}}$, and let $p: \R^3\rightarrow \R^2$ be the projection map onto the first two coordinates (thus $p$ is linear). Then $Q$ is a triangular pyramid with the ``top" vertex skewed. so $|\text{facets}(Q)| = 4$. Clearly $p(Q) = P$ and thus $(Q,p)$ is an extension of $P$ with the same number of facets. But $p$ not a linear isomorphism. Observe that $\frac{1}{3}(1,0,0) + \frac{1}{3}(0,1,1) + \frac{1}{3}(0,0,0) = \frac{1}{3}(1,1,1) \in Q$ and $ \frac{1}{3}(1,1,0) + \frac{2}{3}(0,0,0) = \frac{1}{3}(1,1,0) \in Q$ have the same projection:
$$p(\frac{1}{3}(1,1,1)) = p(\frac{1}{3}(1,1,0)) = \frac{1}{3}(1,1) \in P$$
and so $p$ is not injective and thus not an isomorphism.$\blacksquare$
%Problem 3
\section{}
\paragraph{}
Let $Q \subseteq \R^n$ be a polyhedron and $p:\R^n \rightarrow \R^d$ a linear map, so that $(Q,p)$ is an extension of a polytope $P \subseteq \R^d$. Let $V$ be the set of vertices of $Q$ and $R$ the set of extreme rays of $Q$. That is
$$ Q = \conv{V} + \cone{R}.$$
If $R = \emptyset$ we are done since $Q$ is itself a polytope. So suppose $R\neq \emptyset$. We claim that $p(r) = 0$ for all $r \in R$. Suppose not, that is there exists $r \in R$ for which $p(r) \neq 0$. 
\paragraph{}
Let $v \in V$ if $V\neq \emptyset$ and choose $v = 0$ otherwise. Now, for all $\lambda \geq 0$ $v + \lambda r$ is in $Q$. But then observe that
$$p(v + \lambda r) = p(v) + \lambda p(r)$$
is in $P$. Since $p(r) \neq 0$ this implies that $p(r)$ is an unbounded direction in $P$, contradicting that $P$ is a polytope (and hence bounded). 
%Problem 4
\section{}
\paragraph{}

>>>>

\begin{document}
%Problem 1
\section{}
\paragraph{}
Let a polytope $Q\subseteq \R^n$ and a linear map $p:\R^n \rightarrow \R^d$ be an extension of a polytope $P\subseteq \R^d$. Since $p$ is a linear map, there exists a matrix $A \in \R^{d\times n}$ such that $p(x) = Ax$ for all $x \in Q$.
\paragraph{}
Let $k \in \Z$ such that $0\leq k \leq d$. We will show the existence of an injective map $f$ from the faces of $P$ of dimension $k$ to the faces of $Q$ of dimension at least $k$. This suffices to show the desired result.
\paragraph{}
Let $F$ be a face of $P$ of dimension $k$ such that $F = \{x \in P: c^Tx = \beta\}$. We define the map $f$ so that
$$f(F) = \{x \in Q: (c^TA)x = \beta\}.$$
We need to show that $f$ is well-defined, and that it is injective. To this end we begin by showing that $f(F)$ is a face of $Q$. Let $x \in Q$. Then
$$(c^TA)x = c^T(Ax) \leq \beta$$
with the last inequality following since $p(x) = Ax$ is in $P$. Note that equality holds for all $x$ such that $Ax$ is in $F$. Hence $f(F)$ is a face of $Q$. Since $(Q,p)$ is an extension of $P$, for all $y \in F$ there exists $x \in Q$ such that $Ax = y$. Thus $F \subseteq p(f(F))$, and using that $p$ is a linear map we see that, 
$$\text{dim}(f(F)) \geq \dim F = k.$$
\paragraph{}
Now suppose that $F$ and $F'$ are $k$-dimensional faces of $P$ such that $f(F) = f(F')$. Suppose $F = \{x \in P: c^Tx = \beta\}$ and $F' = \{x \in P: (c')^Tx = \beta\}$. Let $y \in F$. Then there exists $x \in f(F)$ such that $Ax = y$. Since $f(F) = f(F')$,
$$\beta = (c')^TAx = (c')^Ty $$
and hence $y \in F$. Thus $F \subseteq F'$ and by a symmetric argument we can see that $F' \subseteq F$. Thus $F = F'$ and hence $f$ is an injective map, as desired. $\blacksquare$
%Problem 2
\section{}
\paragraph{}
The claim is \emph{false}. Let $P = \conv{\{(0,0),(1,0),(1,1),(0,1)\}}$. Then $P$ is a square, so $|\text{facets}(P)| = 4$. Let $Q = \conv{\{(0,0,0),(1,0,0), (1,1,0), (0,1,1)\}}$, and let $p: \R^3\rightarrow \R^2$ be the projection map onto the first two coordinates (thus $p$ is linear). Then $Q$ is a triangular pyramid with the ``top" vertex skewed. so $|\text{facets}(Q)| = 4$. Clearly $p(Q) = P$ and thus $(Q,p)$ is an extension of $P$ with the same number of facets. But $p$ not a linear isomorphism. Observe that $\frac{1}{3}(1,0,0) + \frac{1}{3}(0,1,1) + \frac{1}{3}(0,0,0) = \frac{1}{3}(1,1,1) \in Q$ and $ \frac{1}{3}(1,1,0) + \frac{2}{3}(0,0,0) = \frac{1}{3}(1,1,0) \in Q$ have the same projection:
$$p(\frac{1}{3}(1,1,1)) = p(\frac{1}{3}(1,1,0)) = \frac{1}{3}(1,1) \in P$$
and so $p$ is not injective and thus not an isomorphism.$\blacksquare$
%Problem 3
\section{}
\paragraph{}
Let $Q \subseteq \R^n$ be a polyhedron and $p:\R^n \rightarrow \R^d$ a linear map, so that $(Q,p)$ is an extension of a polytope $P \subseteq \R^d$. If $Q$ is bounded, we are done, so we may assume that $Q$ is unbounded. Let $R$ be the recession cone of $Q$. We claim that $p(R) = 0$, for otherwise if $r \in R$ with $p(r) \neq 0$ then $p(r)$ is an unbounded direction in $P$ as $p(\lambda r) = \lambda p(r) \neq 0$ for all $\lambda  > 0$, contradicting that $P$ is a polytope. Let $L = R \cap (-R)$ be the lineality space of $Q$. By our previous claim $p(L) = 0$. We have that
$$Q = L + (Q\cap L^\perp)$$
where $L^\perp$ denotes the orthogonal complement of $L$. Thus
$$P = p(Q) = p(L) + p(Q\cap L^\perp) = p(Q\cap L^\perp).$$
Let $Q_1$ denote $Q\cap L^\perp$, so we have shown $p(Q_1) = P$, that is $(Q_1, p)$ is an extension of $p$. Since $Q_1$ was obtained from $Q$ by adding only equalities, $Q_1$ has no more facets than $Q$. Our choice of $Q_1$ ensures that $Q_1$ has a trivial lineality space, i.e. $Q_1$ is pointed. If $Q_1$ is bounded, we are done, so we may assume that $Q_1$ is unbounded. Let $R_1$ be the recession cone of $Q_1$ and $V_1$ be the vertices of $Q_1$. By our prior arguments neither $R_1$ nor $V_1$ are empty.
\paragraph{}
Let $c \in \R^n$ be a vector such that $c^Tr \geq 1$ for all $r \in R_1$. Such a $c$ exists since $R_1 \neq \emptyset$ and the lineality space of $Q_1$ is empty (clearly $c$ so that $c^Tr >0$ exists, and we simply scale $c$). Let $\beta \in \R$ be chosen so that $c^Tv < \beta$ for all $v \in V_1$. Since $V_1$ is finite, such $\beta$ exists. We choose $Q_2$ as
$$Q_2 = Q_1 \cap \{x \in \R^n : c^Tx = \beta\}.$$
We claim that $Q_2$ is bounded and that $p(Q_2) = P$. Let $\alpha = \min\{c^Tv : v \in V_1\}$. Let $x \in Q_2$. There exist $\lambda \geq 0$, $\sum_{v \in V_1} \lambda_v = 1$ and $\gamma \geq 0$ such that
$$x = \sum_{v \in V_1} \lambda_v v + \sum_{r \in R_1} \gamma_r r_1.$$
So
$$\beta = c^Tx =  \sum_{v \in V_1} \lambda_v c^Tv + \sum_{r \in R_1} \gamma_r c^Tr_1 \geq \alpha + \sum_{r \in R_1} \gamma_r c^Tr_1\geq \alpha + \sum_{r \in R_1} \gamma_r$$
and thus $$\sum_{r \in R_1} \gamma_r \leq \beta -\alpha.$$
Hence our choice of coefficients $\gamma$ on the vectors of $R_1$ is bounded, and so $Q_2$ is bounded.
\paragraph{}
It remains to verify that $p(Q_2) = P$. Similarly to the argument for $R$, we see that $p(R_1) = 0$. Hence $P = p(Q_1) = p(\conv{V_1}).$ So it will suffice to show that for every $x \in \conv{V_1}$ there exists $\lambda \geq 0$ and $r \in R_1$ such that $c^T(x + \lambda r) = \beta.$ Let $r$ be any vector in $R_1$ (since $R_1 \neq \emptyset$ this choice is well-defined). Choose $\lambda = (\beta - c^Tx)/c^Tr$. Since $\beta-c^Tx >0$ and $c^Tr >0$ by our choice of $\beta$ and $c$, we have $\lambda \geq 0$. Now we have
$$c^T(x + \lambda r) = c^Tx + \beta - c^Tx = \beta$$
as desired. $\blacksquare$
\section{}
\paragraph{}
